---
title: "Assignment 4"
author: "Gonzalo Del Carpio & Daniel Bies"
date: "9/28/2017"
output: html_document
---

#Data Science with R
##Data tidying
###  http://garrettgman.github.io/tidying/

In the beginning of the article we learn that we are able to organize tabular data in many different ways. We also learn that by default, R makes one layout of data sets much easier to work with than the others.

```{r}
library(DSR)
```


```{r}
table1
```
###Data in tables is easiest to work with if it follows three rules:
1) Each variable has its own column.
2) Each observation is placed in its own row.
3) Each value is placed in its own cell.

In comparison, Raw data is much harder to work with as nothing is sorted out and organized unlike tidy data.

###Tidy Data works well with R because it uses vectorized programming language.
* A data frame is a list of vectors that shows up as a table in R, when the data is tidy all the variables fall into their own vector.


```{r}
table1$cases
```
This command extracts the values of a variable in a tidy data set.


###R makes for an easy user experience for it having the capability to take atomic vectors as input. Making it easy to manipulate data with simple codes.

```{r}
mean(table1$cases)
table1$cases / table1$population * 10000

```

###Tidy data sets, make for much simpler calculations.
For example, if we wanted to see the rate of TB cases per country we would use:

```{r}
table1$cases / table1$population * 10000
```

On the other hand if you don't have a tidy data set it makes the process for a calculation like this more complicated with more steps needed to obtain the value being sought.

###The two most important functions in tidyr are gather(r) and spread()
* These commands alter the tabular data sets while maintaining the values and relationships within the data set.

###Key value pairs are another easy way to record information.
Example:
Ford: 200000 In this pair Ford is the key and 20000 is the value 

###Spread()

The spread() command turns key:value columns into a set of tidy columns.

```{r}
table2
```
And in order to tidy it you would simply use spread(table2, key, value) in order to condense it to tidy data.

###Gather() 

The gather() command does the complete opposite of the spread(), it collects a set of column names and places and places them in a single key column.

###Separate()

This command turns a single character column into multiple columns by splitting the values of the column wherever a separator character appears.

###Unite() 

The unite() command is the exact opposite of the separate() command and combines multiple columns into a single one.


#Tidy Data
##  http://r4ds.had.co.nz/tidy-data.html

# http://r4ds.had.co.nz/transform.html
##Data Transformation Intro
### Dataset Descriptors
* int stands for integers.

* dbl stands for doubles, or real numbers.
* chr stands for character vectors, or strings.
* dttm stands for date-times (a date + a time).
* lgl stands for logical, vectors that contain only TRUE or FALSE.
* fctr stands for factors, which R uses to represent categorical variables with fixed possible values.
* date stands for dates.

### dplyr basics
* Pick observations by their values (filter()).
* Reorder the rows (arrange()).
* Pick variables by their names (select()).
* Create new variables with functions of existing variables (mutate()).
* Collapse many values down to a single summary (summarise()).

### Verb Usage
All verbs work similarly:

1) The first argument is a data frame.
2) The subsequent arguments describe what to do with the data frame, using the variable names (without quotes).
3) The result is a new data frame.

## Filtering Rows With Filter()
#### This allows you to organize observations based off their values

Example:

* filter(dataset, variable = x, 2nd variable = x)

## Narrowing Focus Using Select()
### Allows you to pick the data you want to see
There are a number of helper functions you can use within select():

* starts_with("abc"): matches names that begin with "abc".
* ends_with("xyz"): matches names that end with "xyz".
* contains("ijk"): matches names that contain "ijk".
* matches("(.)\\1"): selects variables that match a regular expression. This one matches any variables that contain repeated characters. You'll learn more about regular expressions in strings.
* num_range("x", 1:3) matches x1, x2 and x3.

## Adding New Variables with Mutate()
### Allows you to add new columns that are functions of existing columns

* Example: mutate(dataset, (newname) = variable (+,-,/,*) variable

## Grouped Summaries with Summarise()
### Used to collapse a dataframe to one, summarized row

Useful Function of summarise():
#### Measure of Location
* mean(x)
* median (x)

#### Measure of Spread
* sd(x) standard deviation
* IQR(x) interquartile range
* mad(x) median absolute deviation

#### Measures of Rank
* min(x)
* quantile(x, 0.25) will return value of x that is greater than 25% of values but less than other 75%
* max(x)

#### Measures of Position
* first()
* nth ()
* last () 




Example:
* by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)

### Ungrouping
* To ugroup data, use the ungroup() function 

# http://rpubs.com/justmarkham/dplyr-tutorial-part-2
## Going Deeper into dplyr
### Choosing Columns
Minus sign can be used to manage columns
* select(-variable, -variable) shows data without those variables
* select(-(variable:variable)) hides a range of columns
* select(-contains("time")) hides columns with matching name
 
## Renaming Columns
### Do not use select()
* Rename(x = y) will rename a variable without deleting all other variables
 
## Advanced Filtering
### Functions to help narrow the scope of data
* filter(between(dep_time, 600, 605)) will return values between 600 & 605
* slice() filters rows by position
 
## Addings New Rows
### Commands to be used when adding to data
* mutate(new variable = (old variables mixed together)) creates a new variable without deleting old ones
* transmute() works like mutate but deletes the other variables
* add_rownames("variable") turns row names into an explicit variable
 
## Grouping and Counting Variables
### Functions to be used to help summarize data
Summarize can be used to count the number of rows in each group:
* Example: group_by(month) %>% summarise(cnt = n())
* tally() and count() perform similar functions
* sum() can be used to add up the total of a variable
 
## Increasing View
### Functions to be used to view more data
* print(n= 15) will specify that you want to see more rows
* view() will show up to 1000 rows and all columns
 